---
title: "HW04"
author: "Forouh Kalantari"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(gapminder) 
library(rmarkdown)
library(tibble)
library(DT)
library(tidyverse)
library(tsibble)
library(scales)
library(dplyr)
library(ggplot2)
library(scales)
library(tidyr)
```

## Exercise 1: Univariate Data Reshaping
### Univariate Option 1
#### Gapminder original dataset 
```{r, echo=TRUE} 
DT::datatable(gapminder)
```

#### 1.1 Make a tibble with one row per year, and columns for life expectancy for Iran, Turkey, Afghanistan.
```{r, echo=TRUE}
gapminder_wide <- gapminder %>%
  filter(country == "Iran"| country == "Turkey" | country == "Afghanistan")%>%
  select(country, year, lifeExp)%>%
  pivot_wider(id_cols = year, 
               names_from = country,
               values_from = lifeExp)
DT::datatable (gapminder_wide) 
```

#### 1.2 scatterplot of life expectancy for Iran (green) vs Afghanistan (red) 
```{r, echo=TRUE}
gapminder_wide %>%
  ggplot( aes(x= year))+ 
  labs(y = "Life Expectancy")+
  geom_point (aes(y = Iran), colour = "orange") +
  geom_point (aes(y = Afghanistan), colour = "black") 

```

#### 1.3 Re-lengthen the data
```{r, echo=TRUE}
 gapminder_wide %>% 
  pivot_longer(cols = -year,
               names_to = "country",
               values_to = "lifeExp") %>%
  DT::datatable()
  
```

## Exercise 2: Multivariate Data Reshaping
### Multivariate Option 1
#### 2.1 make a tibble with one row per year, and columns for life expectancy and GDP per capita for Canada and Germany. 
```{r}
gapminder_wide2 <- gapminder %>%
  filter(country == "Canada"| country == "Germany")%>%
  select(country, year, lifeExp, gdpPercap)%>%
  pivot_wider(id_cols = year, 
               names_from = country,
              names_sep = "_",
               values_from = c(lifeExp, gdpPercap))
DT::datatable (gapminder_wide2) 

```

#### 2.2 Re-lengthen the data
```{r}
gapminder_wide2 %>% 
  pivot_longer(cols = -year,
               names_to = c (".value","country"), 
               names_sep = "_") %>%
  DT::datatable()
  
```


## Exercise 3: Table Joins
## wedding guestlist and email addresses
```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

#### 3.1 For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble. 
