---
title: "Assignment 2 gapminder exploration" 
Author: " Forouh Kalantari"
Date: 2019-09-24
output: html_document
---
###### In this assignment I explore dataset using "gapminder" R package and "tidyverse" Rpackage 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = FALSE)
```

```{r}
library(gapminder) 
library(rmarkdown)
library(tibble)
library(DT)
library(tidyverse)
library(tsibble)
library(scales)
library(dplyr)
library(ggplot2)
```

# gapminder into tibble 

```{r}
datatable(as_tibble(gapminder)) 
```

## Exercise 1: Basic `dplyr`

#1.1 filter the `gapminder` data to include only data from Afghanistan,Canada, and Denmark in the 1970's

```{r}
n <-  gapminder %>%
  filter(country == 'Afghanistan' | country == 'Canada' | country == 'Denmark',
         year >= 1970 & year < 1980)
knitr::kable(n, booktabs = TRUE, align = 'l', caption = "A table of select countries in 1970s ") 

```

#1.2 show the country and GDP per capita for each of the countries filtered.


```{r}
m <- n %>%
  select(country, gdpPercap)
knitr::kable(m, booktabs = TRUE, align = 'l', caption = " Table of country and  gdp ") 
```

#1.3 filter the `gapminder` dataset to all entries that have experienced a drop in life expectancy.


```{r}
gapminder %>%
  group_by(continent, country) %>%
  mutate(lifeExp_inc = lifeExp - lag(lifeExp, order_by = year)) %>%
  filter(lifeExp_inc < 0) %>%
  DT::datatable() %>%
  DT::formatRound(columns = c("lifeExp", "gdpPercap", "lifeExp_inc"),
                  digits = 2)


```

#1.4 filter the `gapminder` dataset so that it shows the max GDP per capita experienced by each country.


```{r}
datatable(gapminder %>%
  group_by(country)%>% 
  summarise(gdpPercap = max(gdpPercap)))
```

#1.5 produce a scatterplot of Canada's life expectancy vs GDP per capita.

```{r}
gapminder %>% 
  filter(country == "Canada") %>%
  select(gdpPercap, lifeExp) %>%
  ggplot(aes(gdpPercap, lifeExp)) + geom_point(alpha= 0.8) + 
scale_x_log10("gdppercap",labels = scales::dollar_format()) +   ylab("life Expectancy") 

```

## Exercise 2: Explore individual variables with `dplyr`

#2.1 pick one categorical variable (continent) and one quantitative variable (gdpPercap). 
# First I plot the data to obtain a visual understanding of data.
# The geom_point plot shows that the Maximum gdpPercap is in Asia and minimum is in africa, also based on data distribution we would expect higher Sd for Asia continent.

```{r}
gapminder %>% 
  select(continent,gdpPercap) %>%
  mutate(continent= factor(continent)) %>% 
  ggplot(aes(continent, gdpPercap)) +
  scale_y_log10("gdpPercap", labels=scales::dollar_format()) +
  geom_point() + xlab("continent")
```


#2.2The table below reflects the data from geom_point plot. 

```{r}
gapminder %>% 
  select(continent,gdpPercap)%>%
  group_by(continent)%>%
  summarise(max_gdp = max(gdpPercap), min_gdp = min(gdpPercap), 
            mean_gdp = mean(gdpPercap),median_gdp = median(gdpPercap),sd_gdp = sd(gdpPercap))
```


## Exercise 3 Explore various plot types

#This is a scatterplot of continents life expectancy vs GDP per capita. 

```{r}
 gapminder %>%
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point(aes(color = continent)) +
  scale_y_log10() +
  labs(x = "Life Expectancy",
       y = "gdpPercap",
       color = "Continent")
```

#3 2nd plot.is is a boxplot of  the life expectancy of continents between 2003-2007
```{r}
gapminder %>%
  filter(year >= 2003) %>%
  ggplot(aes(continent, lifeExp)) +
  geom_boxplot() +
  labs(x = "Continent",
       y = "Life Expectancy (years)")
```
