---
title: "HW05"
author: "Forouh Kalantari"
output: html_document
---

```{r,r, warning=FALSE, message=FALSE}
library(gapminder)
library(tidyverse)
library(dplyr)
library(forcats)
library(ggplot2)
library(tibble)
library(here)
library(DT)
library(ggridges)
library(scales)
library(gridExtra)

```
## Exercise 1: Explain the value of the here::here package
#### here::here package is specifically useful when diffirent people are collaborating in R. In General, it will provide a local copy of the project for the collaborators. here::here will be functional by specifying only the path from the project root, regardless of where the local copy of the project is saved for each collaborators. It will also make it easier to detect the root directory used in a R studio project. 

## Exercise2: factor management

#### In the gapminder data set I explore continent variable. first, I confirm that continent is a factor.
```{r}
gapminder$continent %>%
  class()
```

#### Then I check to see how many levels are there in this factor
```{r}
gapminder$continent %>%
  nlevels()
```

#### Then I identify these different levels which are "Africa"   "Americas" "Asia"     "Europe"   "Oceania"
```{r}
gapminder$continent %>%
  levels()
```

#### I get the number of entries per level and the table below shows it. 
```{r}
gapminder %>%
  count(continent) %>%
  datatable()
```

#### Now I drop Ocenia from list and it is no longer in levels. The following table confirms it. 
```{r}
drop_ocn = gapminder %>%
  filter(continent != "Oceania") %>%
  mutate(continent = fct_drop(continent))
  levels(drop_ocn$continent)
  
 drop_ocn %>%
  count(continent) %>%
  datatable()
```

### Reorder the levels of country or continent. Use the forcats package to change the order of the factor levels, based on summarized information of one of the quantitative variables. Consider experimenting with a summary statistic beyond the most basic choice of the mean/median. Use the forcats package in the tidyverse for this, rather than the baseR function as.factor

```{r}
gap_reorder = gapminder %>%
  mutate(continent = fct_reorder(continent, gdpPercap, max))

levels(gap_reorder$continent)
```

#### Based on this we see that the level of continent that was 'Africa' Americas' ' Asia''Europe'and 'Oceania' has changed to "Africa"   "Oceania"  "Americas" "Europe"   "Asia" (in an order from min to max). 
```{r}
gapminder%>%
  group_by(continent, gdpPercap) %>%
  summarise(maxgdp = max(gdpPercap)) %>%
  ggplot(aes(continent, gdpPercap, color = continent)) +
  geom_point(alpha = 1) +
  ylab("gdppercap")

gap_reorder%>% 
  group_by(continent, gdpPercap) %>%
  summarise(maxgdp = max(gdpPercap)) %>%
  ggplot(aes(continent, gdpPercap, color = continent)) +
  geom_point(alpha = 1) +
  ylab("gdppercap")
```


### Explore the effects of re-leveling a factor in a tibble. I re-level the data that I have the levels'Europe' and 'Americas' are first. 
```{r}
gap_relevel = gapminder %>%
  mutate(continent = fct_relevel(continent,"Europe", "Americas"))

levels(gap_relevel$continent)
```


### Based on this we can see the difference in data tables 
```{r}
gap_reorder %>%
  arrange(continent) %>%
  DT::datatable()

gap_relevel %>%
  arrange(continent) %>%
  DT::datatable()


```

### Also we can see the difference ploting releveled and reordered one. 
```{r}
gap_reorder%>% 
  group_by(continent, gdpPercap) %>%
  summarise(maxgdp = max(gdpPercap)) %>%
  ggplot(aes(continent, gdpPercap, color = continent)) +
  geom_boxplot() +
  theme_bw() +
  ylab("gdppercap")

gap_relevel%>% 
  group_by(continent, gdpPercap) %>%
  summarise(maxgdp = max(gdpPercap)) %>%
  ggplot(aes(continent, gdpPercap, color = continent)) +
  geom_boxplot() +
  theme_bw() +
  ylab("gdppercap")

```

## Exercise 3: File input/output

```{r}
gap_new <- gapminder %>%
  filter(year == max(year)) %>%
  group_by(continent, year) %>%
  summarize(meanlifeExp = mean(lifeExp)) %>%
  pivot_wider(id_cols = continent,
              names_from = year,
              names_prefix = "mean lifeExp year",
              values_from = meanlifeExp)
  
gap_new %>%
  DT::datatable()
```



```{r}
write_csv(gap_new,
          here("hw05", "data", "gapminder_new_meanlifeExp.csv"))
```

```{r}
gap_new_csv = read_csv(here("hw05", "data", "gapminder_new_meanlifeExp.csv"))

gap_new_csv %>%
  DT::datatable()
```

```{r}
gap_new_csv$continent %>%
  class()

gap_new$continent %>%
  class()


```



## Exercise 4: Visualization design. Create a side-by-side plot and juxtapose your first attempt (show the original figure as-is) with a revised attempt after some time spent working on it and implementing principles of effective plotting principles. Comment and reflect on the differences.
#### I will re plot Hw03 (Task Option 3) the spread of GDP per capita within the continents
#### This is the original plot I made in that assignment. 

 
```{r}
# before 
old_plot = gapminder %>% 
  ggplot(aes(gdpPercap, continent, fill = continent)) + 
  ggridges::geom_density_ridges() 
old_plot
```

#### 1-changing the x-axis to a log scale so that the spread of all  continents are not towards left. There is a different pattern of spread in all continents for GdpPercap now.  
#### 2-relabeling the plot so that we know what is the purpose of this plot
#### 4- Being able to define a color of ineterest
#### 5-replacing the grey theme with white so that it is more visible 
```{r}
#after
new_plot = gapminder%>%
ggplot(aes(gdpPercap, continent, fill = continent)) +
  geom_density_ridges(alpha = 0.7) +
  scale_discrete_manual("fill", values = c("green", "red", "purple", "blue", "orange")) +
  scale_x_log10(labels = comma_format()) +
  labs(title = "Spread of gdppercap per Continent",
       x = "GdpPercap",
       y = "Continent") +
  theme_bw() +
  theme(legend.position = "none")
new_plot

```

#### we can compare 2 plots side-by-side (Old plot vs to new plot)
```{r}
grid.arrange(old_plot,new_plot, nrow = 1)
```


## Exercise 5: Writing figures to file

#### I am saving the newplot form Exercise 4 into plot file in my Hw05 folder.
```{r}
ggsave(here("hw05", "plot", "EX4_newplot.png"),
       plot = new_plot,
       width = 3, height = 2.5, scale = 1.5)

```
