---
title: "Untitled"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(gapminder) 
library(rmarkdown)
library(tibble)
library(DT)
library(tidyverse)
library(tsibble)
library(scales)
library(dplyr)
library(ggplot2)
```


#(Task Option 2)The maximum and minimum of GDP per capita for all continents.

#The table below and ggplot show maximum and minimun of gdpPercap for all continents.

```{r}
gapminder %>% 
group_by(continent) %>% 
summarize(maxgdp = max(gdpPercap),mingdp= min(gdpPercap))
```



```{r}
gapminder %>%

  ggplot(aes(continent, gdpPercap)) +

  geom_boxplot(size = 1.5) +

  scale_y_log10(labels = scales::dollar_format()) +

  labs(title = "GDP per Capita per Continent", x = "Continent", y = "GDP per Capita")
```



#(Task Option 3)Look at the spread of GDP per capita within the continents.

```{r}
# measures of spread = range (max- min), Inter-quartile range, standard deviation, median absolute deviation

a1 <- gapminder %>%
  filter(continent == "Africa") %>%
  group_by(year) %>%
  summarise(range= max(gdpPercap) - min(gdpPercap),
            IQR = IQR(gdpPercap),
            Standard_Deviation = sd(gdpPercap),
            MAD = mad(gdpPercap))

a1 %>%  ggplot(aes(year,range))+
  geom_line(alpha = 0.5)+
  xlab("Year") + ylab("Range")



gapminder %>% 
  ggplot(aes(gdpPercap, continent, fill = continent)) +
  ggridges::geom_density_ridges()

```





#(Task Option 5)How is life expectancy changing over time on different continents?


```{r}
gapminder %>%
ggplot(aes(year, lifeExp, group = country, colour = continent)) +
  geom_line(alpha = 0.5) +
  scale_x_log10( labels = scales::comma_format())+
  facet_wrap(~ continent)

Africa <- gapminder %>%
  filter(continent == "Africa") %>%
  group_by(year) %>%
  summarise(mean= mean(lifeExp))

Africa
  


```

#Based on this plot except for some countries in Africa and a few countries in Asia and America life expectancy has increased for most countries over the years. 